---
title: "Learning the Tidyverse"
author: "katherine_powelson@fws.gov"
date: "11/6/2019"
output: html_document
---
Primary key is something that does not repeat, a foreign key repeat. Something could be a primary key in one table and a foreign key in another table. 
git remove -v in terminal can make sure that you are in the right folder/place

ctrl+sft+m makes a pipe  it takes the output from a output and uses it for the input of the next . Pipe is like the work then
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(dplyr)
library(tidyr)
#package_name::function_name(...) you can uses this to unmask functions  
```
# Read and clean data 
```{r}
catch_orginal <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",method="libcurl"),stringsAsFactors = F)

head(catch_orginal)
```


# Reshape our data 
 We are going to drop the notes and the all colum. For Tidyverse .data is always the first argument, it automatically populates with the product of previous function
```{r}


catch_data <- catch_orginal %>% 
  #select(Region, Year,Chinook,Sockeye,Coho,Pink,Chum)#select function chooses and then renames colums from a table.This tells what to keep. 
  select(-All, -notesRegCode)# this way tells you which colums to get rid of 
head(catch_data)
```
Chinook is a chr now we need to mutate it 
```{r}
catch_clean <-catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```
Now we will find the rows that got turned into NA my the mutate function 
```{r}
(i<-which(is.na(catch_clean$Chinook)))
```
lets print from catch_orginal the [] let you look at a colum by index number 
```{r}
catch_orginal[i,]
```
Now lets fix this with out mutate
```{r}
catch_clean <-catch_data %>% 
  mutate(Chinook = ifelse(Chinook=="I",1,Chinook)) %>% 
  mutate(Chinook=as.numeric(Chinook))
summary(catch_clean)

```
# Reshape our data 
Pivot longer 
```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region,Year),# just like above you can you all c() or -c()
  names_to="species",# the name for the pivot value
  values_to= "count")
```
This is the opposite of pivot longer, pivot wider. This is needed for making tables and stuff 
```{r}
catch_wide <- catch_long %>%
  pivot_wider(names_from= Year, #makes a colum for each year
              values_from = count)

catch_long <-  catch_long %>% 
  rename(catch_thousands =count)

head(catch_long)  

```
```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousands*1000) %>% 
  select(-catch_thousands)
head(catch_long)
```

# Summarize data
This summarizes by similar values
```{r}
mean_region <- catch_long %>% 
  group_by(Region, species) %>% #group by
  summarize(catch_mean=mean(catch),#tell it what to do with each group
            num_obs=n())# this tells you the number of observations 
mean_region

catch_by_spp <- catch_long %>% 
  group_by(species) %>% 
  summarize(total_catch=sum(catch)) %>% 
  arrange(desc(total_catch))#arrangeing the values

catch_by_spp
```

filter 

```{r}
catch_by_spp_pre1900 <- catch_long %>% 
  filter(Year<1990) %>% 
  group_by(species) %>% 
  summarize(total_catch=sum(catch)) %>% 
  arrange(desc(total_catch))
  
catch_by_spp_pre1900
```

# Joins to region table 
## Adding colum that will be joined 
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method="libcurl"),
                        stringsAsFactors = F) %>% 
  select(code, mgmtArea)# the columns that we need

```


```{r}
#catch_joined <- left_join(catch_long, region_defs)# it is trying go guess what to join by but it can't because they are named differently 
```

 two solutions either specify what to join by
```{r}
catch_joined <- left_join(catch_long, region_defs, by= c("Region"="code")) 

```
or  rename it when you are bring it in 

# Seperate and Unite
```{r}
site_codes <- data.frame(site = c("HAW-100",
                                  "HAW-101",
                                  "OAH-102",
                                  "MAI-100"),
                         stringsAsFactors = F)
```
seperate Island and site number 

```{r}
site_codes_split <- site_codes %>% 
  separate(site, into =c("island","site_number"), sep="-")# this is the orginal colum "site" and the name for the new colums 
site_codes_split
```
gsub package will look for a pattern instead of the symbol 

Now we will recombine them 

```{r}
united <- site_codes_split %>% 
  unite(col= site_code, island, site_number, sep="_")

united
```



