---
title: "Spatial_data"
author: "katherine_powelson@fws.gov"
date: "11/8/2019"
output: html_document
---
datum = shape of the sphere 
CRS = is just a reference system
area and distance measurements require a projection
Projection is not always needed 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
library(leaflet)
```

# Prepare the regions shapefile 
```{r}
ak_regions <-read_sf("data/shapefiles/data/shapefile_demo_data/ak_regions_simp.shp")
st_crs(ak_regions)
class(ak_regions)
ak_regions_3338 <- ak_regions %>% 
  st_transform(crs=3338)

plot(ak_regions_3338)

summary(ak_regions_3338)
```
```{r}
summary(ak_regions_3338)

ak_regions_3338 %>% 
  select(mgmt_area)
```

# Prepare the population by region 
```{r}
pop <-read.csv("data/shapefiles/data/shapefile_demo_data/alaska_population.csv",stringsAsFactors = F)

class(pop)

pop_4326 <- st_as_sf(pop,
                     coords =c("lng","lat"),
                     crs = 4326,
                     remove =F)# pass the crs that is currently is, not the one you want

pop_3338 <- st_transform(pop_4326,crs=3338)
#pop_3338<- st_coordinates() #add coordinate,
#add a colum with the ESPG

pop_joined <- st_join(pop_3338, ak_regions_3338, join=st_within)# we want to know which of the points is within each region
head(pop_joined)

```
```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop =sum(population))

head(pop_region)
```
Now we need to get it back to the correct geometries. 

```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region)
# you could specify, which you would need if the colums did not have the same header
#pop_region_3338 <- left_join(ak_regions_3338, pop_region, by=c("region"="region")


```


```{r}
#dissolve the boundaries 
pop_mgmt <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop=sum(total_pop))

plot(pop_mgmt["total_pop"])

# to keep internal boundaries 

pop_mgmt <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop=sum(total_pop), do_union=F)


plot(pop_mgmt["total_pop"])
pop_mgmt$ESPG = "3338"
?sf::tidyverse
```

# Make maps in ggolot! 

```{r}

ggplot()+
  geom_sf(data = pop_region_3338, mapping = aes(fill=total_pop))+
  #geom_sf(data = rivers_3338, aes(size = StrOrder), color = "black") +
  geom_sf(data = pop_3338, aes(), size = .5) +
  scale_size(range = c(0.01, 0.2), guide = F) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```
# Adding base maps in ggmap 

3857 pseudomercador is the project on the tiels 

```{r}
pop_3857 <- pop_3338 %>% 
  st_transform(crs =3857)

# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```


```{r}
bbox <- c(-170, 52, -130, 64)
ak_map <- get_stamenmap(bbox,zoom = 4)
ak_map_3857 <- ggmap_bbox_to_3857(ak_map)
```

```{r}
ggmap(ak_map_3857)+
  geom_sf(data=pop_3857, aes(color=population),inherit.aes=F)+
  scale_color_continuous(low="khaki", high ="firebrick", label=comma)

```

# Leaflet 
leaflet will project your files for you but you need to make sure you use somehting lik wgs 84 (4326)
in needs to be a crs with lat long 
```{r}
# this would be a good function 
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

```{r}
pop_region_4326 <- pop_region_3338 %>% 
  st_transform(crs=4326)
```

```{r}
leaflet(options=leafletOptions(crs= epsg3338)) %>% 
  addPolygons(data=pop_region_4326,
              fillColor = "gray",
              weight =1)

#add legend scale 
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),# tells how to map it to the pallette created above
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
        addLegend(position = "bottomleft", #adds a legend to the map
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")
m
```
```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1) %>% 
        addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```

