---
title: "data_visualization"
author: "katherine_powelson@fws.gov"
date: "11/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F, warning=F}
library(ggplot2)
library(leaflet)
library(dplyr)
library(tidyr)
library(scales)
library(DT)
library(lubridate)
```

# Prepare data
Try catch trys to read the error and then if doesn't work it gives you an alternate. 
This is good because if you are sharing code it will look for the file locally 
```{r}
#data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        #add an if the folder doesn't exist make it 
        return(esc)
    }
)

head(esc)
```

Calculate the Annual escapment by speceis, region, and Year 
  *only for Sockeye, Chinook, Coho, Pink, Chum 
  
```{r}

salmon_data <- esc %>% 
  filter(Species %in% c("Sockeye","Chinook", "Coho")) %>% 
  mutate(sampleDate= as.Date(sampleDate))%>% 
  mutate(Year=year(sampleDate)) %>% 
  group_by(Species,SASAP.Region,Year) %>% 
  summarize(totalesc=sum(DailyCount))

# without lubridate 
salmon_data <- esc %>% 
  filter(Species %in% c("Sockeye","Chinook", "Coho","Pink","Chum")) %>% ## you can us == in 
  separate(sampleDate, into = c("Year", "Month", "Day")) %>% 
  mutate(Year=as.numeric(Year)) %>% 
  group_by(Species,SASAP.Region,Year) %>% 
  summarize(total_esc=sum(DailyCount))

head(salmon_data)
```

# Create static plots 

```{r}
ggplot(salmon_data,aes(x=Species, y=total_esc, fill=SASAP.Region))+
  geom_col()

```
```{r}
kodiak_esc <- salmon_data %>% 
  filter(SASAP.Region=="Kodiak")

my_theme <- theme_bw()+
  theme(legend.position = "bottom")
pal <- c("blue", "green", "khaki", "gray", "yellow")
ggplot(kodiak_esc,aes(x=Year, y=total_esc,color=Species))+
  geom_line()+
  geom_point()+
  scale_color_manual(values=pal)+
  scale_y_continuous(labels = comma)+
  theme_bw()+ # call default theme before other theme edits
  theme(legend.position = "bottom")

```


# Create an interactive table 

# Create and interactive map

```{r, fig.height=7, fig.width=7}
ggplot(salmon_data, aes(x=Year, y=total_esc, color=Species))+
  geom_line()+
  geom_point()+
  facet_wrap(~SASAP.Region, ncol=2, scales = "free")+
  scale_y_continuous(labels = comma)
  
```


```{r}
locations <- esc %>% 
  distinct(Location, Latitude,Longitude) %>% #pulls out distinct combinatiosn
  drop_na()

datatable(locations)
```
#Leaflet
leaflet uses pipes instead of the leaflet is WGS84
```{r}
leaflet(locations) %>% 
  addTiles() %>% #add the default tiles(basemap)
  addMarkers(lng=~Longitude, lat = ~Latitude, popup=~Location)
  
```

```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```



